---
editor_options:
  markdown:
    wrap: 72
output: word_document
---

# [DS202_Final]{.smallcaps}

Repo for DS202 final project

```{r}
library(WDI)
library(dplyr)
library(tidyverse)
library(ggthemes)

```

# [Changes in CO2 Emissions Over Time Members: Katherine Amundson, Kendra Scace, Gautham Suresh, Kelly Ryan]{.smallcaps}

## [[**Introduction**]{.underline}]{.smallcaps}

Understanding the historical trends of CO2 emissions is essential for
crafting policies and strategies to mitigate their impact. In this
project, we explore the changes in CO2 emissions overtime, focusing on
different aspects such as the comparison of emission patterns between
certain countries, impact of international agreements, and specific
countries with the highest emissions. Along with that, we showed
comparisons of CO2 emissions over time and the difference in emissions
between developed and developing countries.

## ~[[**Data**]{.underline}]{.smallcaps}~

We utilized a comprehensive dataset on CO2 emissions by country from
1750 to 2022. The dataset includes information on each country's
emissions, population, area, and relevant codes. After loading the data,
we conducted data cleaning to handle missing values and ensure the
dataset's integrity.

```{r}
df <- read.csv("CO2 emission by countries.csv")
head(df)
length(df)
nrow(df)
summary(df)
```

The dataset contains information about CO2 emissions from different
countries over a significant period, from 1750 to 2020. The dataset
comprises 9 columns:

1.  **Country**: Represents the name of the country.

2.  **Code**: Corresponds to the country's abbreviation or code.

3.  **Calling Code**: Refers to the area code in a phone number
    associated with the country.

4.  **Year**: Denotes the year when the data was observed.

5.  **CO2 Emission (Tons)**: Represents the recorded amount of CO2
    emissions by the country for a specific year.

6.  **Population as of 2022**: Indicates the number of people living in
    the country as of the year 2022.

7.  **Area**: Represents the size of the country in square kilometers.

8.  **Percent of World**: Indicates the proportion of the Earth's
    surface area occupied by the country.

9.  **Density in km2**: Represents the population density of the country
    calculated as the number of people per square kilometer.

### [**Dataset Overview:**]{.smallcaps}

-   The dataset contains 59,620 rows, each representing a unique
    combination of country, year, and associated data.

-   The years covered in the dataset range from 1750 to 2020.

-   The CO2 emissions column reveals a wide range of values, from 0 to
    4.17e+11 (4.17e11 represents 4.17 multiplied by 10 to the power of
    11).

-   The population column shows variations in population size, with a
    minimum of 11,310 and a maximum of 1.426e+09 (1.426e09 represents
    1.426 multiplied by 10 to the power of 9).

-   The dataset also includes information about the size of each
    country, its share of the world, and population density.

```{r}
df <- drop_na(df)
```

### [Data Exploration and Cleaning]{.smallcaps}

```{r}
# checking for missing values
sum(is.na(df))
# removing rows with any missing values
df_clean <- na.omit(df)
#provides the structure of the dataframe, showing data types and the first few values of each variable, used for understanding the data types of each column
print("Structure: ")
str(df)
# returns the names of the columns in the dataframe
print("Column names: ")
colnames(df)
#returns the class of the object
print("class: ")
sapply(df, class)
# returns the number of unique values for each variable in the dataframe. Helps with understanding the cardinality of each variable
print("unique values: ")
sapply(df, function(x) length(unique(x)))
# remove duplicate rows
df_unique <- unique(df)
# computes the correlation matrix for the numeric variables in the dataframe. Helps with understanding the relationships between different variables
print("correlation matrix: ")
numeric_df <- df[sapply(df, is.numeric)]
cor_matrix <- cor(numeric_df)
print(cor_matrix)
```

**Checking for missing values:** The result indicates that there are
11,111 missing values in the dataset

**Removing Rows with Missing Values:** clean the dataset by removing
rows with any missing values

**Understanding the Data Structure:** The structure of the dataframe,
including data types and the first few values of each variable, is
printed.

**Column Names and Classes:** The names and classes of the columns in
the dataframe are printed. The columns include "Country," "Code,"
"Year," "CO2.emission..Tons.," "Population.2022.," "Area,"
"X..of.World," and "Density.km2.". Classes range from character to
integer and numeric.

**Unique Values and Cardinality:** The number of unique values for each
variable in the dataframe is printed. The output provides insights into
the cardinality of each variable.

**Correlation Matrix:** The correlation matrix helps understand
relationships between different numeric variables. In this case, there
is a 13.02% positive correlation between "Year" and
"CO2.emission..Tons.".

**Questions to be addressed**

-   Are currently developing countries polluting at rates similar to
    countries that industrialized earlier
-   Have developed countries taken strides towards reducing pollution in
    the past decade?
-   Have international agreements made any observable impact on
    emissions for participating countries?

## [Exploring Unique Countries and Minimum CO2 Emission Years]{.smallcaps}

### [Unique Countries:]{.smallcaps}

The unique countries present in the dataset are extracted and the head is 
printed.

This provides a list of distinct countries represented in the dataset.

```{r}
unique_countries <- unique(df$Country)

head(unique_countries)
```

### [Filter Non-Zero CO2 Emissions:]{.smallcaps}

This step involves excluding entries with zero CO2 emissions, focusing
on non-zero emission instances. This filtered dataset has 21,707 rows
and 9 columns.

```{r}
df <- df %>%
  filter(CO2.emission..Tons. != 0)
```

### [Determining Minimum CO2 Emission Years:]{.smallcaps}

For each country, the year with the minimum CO2 emissions is determined,
and the corresponding data is summarized. This results in a dataset that 
provides insights into the earliest year with non-zero CO2 emissions for each
country, for which we printed the head.

```{r}
min_year_data <- df %>%
  group_by(Country) %>%
  filter(CO2.emission..Tons. != 0) %>%
  summarise(Min_Year = min(Year),
            CO2_Emission = CO2.emission..Tons.[which.min(Year)])

head(min_year_data)
```

## [[**Methods**]{.underline}]{.smallcaps}

Our analysis involved various methods, including data cleaning,
exploratory data analysis, and visualizations. The methods can be
categorized into data handling, exploratory data analysis (EDA), and
visualization techniques. The primary tools utilized for these tasks
were the **`dplyr`** and **`ggplot2`** packages in R.

Our methodology encompassed thorough data cleaning, exploratory data
analysis, and diverse visualizations to uncover patterns and trends in
CO2 emissions across countries, with a specific focus on the impact of
international agreements, top emitters, and the comparison between
countries like China and the United States. The use of R packages,
particularly **`dplyr`** and **`ggplot2`**, facilitated efficient data
manipulation and visualization

## [[**Results**]{.underline}]{.smallcaps}

### [Changes Over Time]{.smallcaps}

We started by examining how emissions have evolved over time for
different countries. For countries with data available before 1900, we
visualized their emission trends. This helped us identify early
industrializers and compare their patterns with currently developing
countries.

```{r}

# Filter countries with Min_Year < 1900
selected_countries <- min_year_data %>%
  filter(Min_Year < 1900)

# Filter data for selected countries
selected_df <- df %>%
  filter(Country %in% selected_countries$Country)

# Get the top 10 countries based on some criterion (e.g., average CO2 emissions)
top_10_countries <- selected_df %>%
  group_by(Country) %>%
  summarize(avg_emission = mean(CO2.emission..Tons.)) %>%
  top_n(10, wt = avg_emission) %>%
  pull(Country)

# Filter the data for the top 10 countries
selected_df_top10 <- selected_df %>%
  filter(Country %in% top_10_countries)

# Plotting
ggplot(selected_df_top10, aes(x = Year, y = CO2.emission..Tons., color = Country)) +
  geom_line() +
  labs(title = "Top 10 CO2 Emissions of Countries with Min Year < 1900",
       x = "Year",
       y = "CO2 Emission (Tons)") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)

```

```{r}
population_threshold <- 75000000  # 75 million

large_countries <- df %>%
  filter(Population.2022. > population_threshold)
```

## [Impact of International Agreements]{.smallcaps}

We investigated whether international agreements have influenced CO2
emissions by examining emission trends for countries with populations
exceeding 75 million. The visualizations included vertical lines
representing the years of key international agreements. The Kyoto
Protocol, created in 1997, and the Paris Agreement, enacted in 2015, are
two of the most influential climate agreements to date, so we chose to
use those as our markers.

### [CO2 Emissions Compared to Dates of International Agreements]{.smallcaps}

```{r}
countries_agreements <- large_countries %>%
  filter(Year > 1985)

head(countries_agreements)

agreements_years <- ggplot(countries_agreements, aes(x = Year, y = CO2.emission..Tons., color = Country)) +
  geom_line() +
  labs(title = "CO2 Emissions Compared to Dates of International Agreements",
       x = "Year",
       y = "CO2 Emission (Tons)") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)
agreements_years + geom_vline(xintercept=1997)
agreements_years + geom_vline(xintercept=2015)
```

### [Highest Emissions in 2020]{.smallcaps}

We identified the top emitters in 2020 and visualized their emissions.
This analysis provides insights into which countries contribute most to
global emissions.

```{r}
largest_emissions <- df%>% 
  filter(Year == 2020) %>%
  slice_max(CO2.emission..Tons., n = 10)

print(largest_emissions)

most_emissions_graph<-ggplot(largest_emissions, aes(x=Country, y=CO2.emission..Tons., col=Country)) +
  geom_point(size = 5) +
    labs(title = "Top 10 Countries by CO2 Emissions in 2020",
       x = "Country",
       y = "CO2 Emission (Tons)") +
  theme_minimal() +
    theme(axis.text.x = element_text(angle = 90)) + scale_y_continuous(labels = scales::comma)



print(most_emissions_graph)
```

Next, we decided to compare emissions from countries with the highest populations.
This would allow us to compare data and see whether countries with high populations
also tend to create more CO2 emissions.

```{r}
emissions_2020 <- df %>%
  filter(Year==2020) %>%
  slice_max(Population.2022., n = 20)

highest_population_emissions<-ggplot(emissions_2020, aes(x=Country, y=CO2.emission..Tons., col=Country)) +
  geom_point(size = 5) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1)) + scale_y_continuous(labels = scales::comma)
    labs(title = "Emissions of the Top 20 Most Populated Countries in 2020",
       x = "Country",
       y = "CO2 Emission (Tons)")

print(highest_population_emissions)
```

### [Global Distribution]{.smallcaps}

To understand the geographical distribution of emissions in 2020, we
created a world map highlighting each country's contribution.

```{r}
# Load world map data
world_map <- map_data("world")
# Rename regions for consistency
world_map <- world_map %>%
  mutate(region = case_when(
    region == "UK"  ~ "United Kingdom",
    region == "USA" ~ "United States",
    TRUE ~ region
  ))
# Filter the data for the year 2020
df_2020 <- df %>% filter(Year == 2020)
# Perform inner join with the world map data
merged_data <- inner_join(world_map, df_2020, by = c("region" = "Country"))
# Plot the global distribution of CO2 emissions for the year 2020
ggplot(merged_data, aes(x = long, y = lat, group = group, fill = CO2.emission..Tons.)) +
  geom_polygon(color = "white") +
  labs(title = "Global Distribution of CO2 Emissions (Year 2020)",
       fill = "CO2 Emissions (Tons)") +
  theme_map()


```

*The map provides a visual representation of the global distribution of
CO2 emissions, allowing for the identification of regions or countries
with notable emission levels. Patterns and concentrations of high or low
emissions can be observed, providing insights into the geographical
variations of carbon dioxide output*

```{r}
valid_country_codes <- df %>%
  distinct(Code) %>%
  pull(Code)


wdi_data <- WDI(country = valid_country_codes, indicator =c("SP.POP.TOTL", "NY.GDP.MKTP.KD"))

```

```{r}
wdi_data <- left_join(df, wdi_data, by = c("Code" = "iso2c", "Year" = "year"))

# WDI only contains population data from 1960
wdi_data <- wdi_data %>%
  filter(Year >= 1960)
```

### [CO2 Emissions Comparison: China vs. United States]{.smallcaps}

```{r}
china_us_data <- wdi_data %>%
  filter(Country %in% c("China", "United States"))

# Plot the pollution rates for China and the US
ggplot(china_us_data, aes(x = Year, y = CO2.emission..Tons., color = Country)) +
  geom_line() +
  labs(title = "CO2 Emissions Comparison: China vs. United States",
       x = "Year",
       y = "CO2 Emissions (Tons)",
       color = "Country") + scale_y_continuous(labels = scales::comma)
```

*Comparing CO2 US and China for the past 60 years shows a general trend
where the emissions rate for China is climbing faster than the US. This
lines up with historical context as companies tend to move manufacturing
plants to countries with looser labor laws and restrictions.*

### [Factors Influencing Emissions in the United States]{.smallcaps}

We explored potential factors influencing emissions in the United
States, including population size and GDP.

```{r}
us_data <- wdi_data %>%
  filter(Country == "United States")

# Plot population vs. emissions rate for the United States
ggplot(us_data, aes(x = SP.POP.TOTL, y = CO2.emission..Tons.)) +
  geom_point() +
  labs(title = "Population vs. Emissions: United States",
       x = "Population",
       y = "Emissions")

ggplot(us_data, aes(x = SP.POP.TOTL, y = CO2.emission..Tons. / SP.POP.TOTL)) +
  geom_point() +
  labs(title = "Population vs. Emissions: United States",
       x = "Population",
       y = "Emissions Rate (Tons per capita)")
```

```{r}
# Filter data for the United States
us_data <- wdi_data %>%
  filter(Country == "United States")

# Create a new variable for faceting
facet_var <- rep(c("Total", "Per Capita"), length.out = nrow(us_data))

# Add the facet_var variable to the dataset
us_data <- mutate(us_data, facet_var = facet_var)

# Plot population vs. emissions rate for the United States
ggplot(us_data, aes(x = SP.POP.TOTL, y = CO2.emission..Tons.)) +
  geom_point() +
  labs(title = "Population vs. Emissions Rate: United States",
       x = "Population",
       y = "Emissions Rate") +
  facet_wrap(~facet_var, scales = "free_y", ncol = 1) +
  ylab(NULL) +
  theme(strip.background = element_blank(), strip.text.x = element_blank()) + scale_y_continuous(labels = scales::comma)

```

```{r}
# Filter data for the United States
us_data <- wdi_data %>%
  filter(Country == "United States")

# Create a new variable for faceting
facet_var <- rep(c("Total", "Per Capita"), length.out = nrow(us_data))

# Add the facet_var variable to the dataset
us_data <- mutate(us_data, facet_var = facet_var)

# Plot population vs. emissions rate for the United States
ggplot(us_data, aes(x = SP.POP.TOTL, y = CO2.emission..Tons.)) +
  geom_point() +
  labs(title = "Population vs. Emissions Rate: United States",
       x = "Population") +
  facet_grid(facet_var ~ ., scales = "free_y", switch = 'y') +
  ylab("Emissions Rate (Total)") +
  theme(strip.background = element_blank(), strip.text.x = element_blank()) +
  scale_y_continuous(labels = scales::comma) +
  theme(strip.background = element_blank(), strip.text.x = element_blank())

```

### [GDP vs. CO2 Emissions: United States]{.smallcaps}

```{r}
ggplot(us_data, aes(x = NY.GDP.MKTP.KD, y = CO2.emission..Tons.)) +
  geom_point() +
  labs(title = "GDP vs. CO2 Emissions: United States",
       x = "GDP",
       y = "CO2 Emissions (Tons)")
```

*This graph higlights a trend where an increase in GDP contributes to
increased C02 emissions*

### CO2 Emissions Comparison

In this section, we delve into a detailed analysis of CO2 emissions
data, focusing on the year 2020. The dataset under consideration
includes information on CO2 emissions, population figures, and
additional pertinent variables across various countries.

1.  Data Preprocessing:

Column Renaming: For enhanced readability and ease of use, columns were
renamed. The column "CO2.emission..Tons." was abbreviated to "coTons,"
while "Population.2022." became "population."

Data Sorting: To facilitate subsequent analyses, the dataset was sorted
in descending order based on population.

2.  Analysis for the Year 2020:

Filtering Data: The dataset was filtered to exclusively include entries
where the "Year" column corresponds to 2020.

Handling NaN Values: A subset of data was extracted where at least one
value had NaN on all columns.

```{r}

# renaming so its easier to type
df1 <- df %>%
  rename(coTons = "CO2.emission..Tons.", population = "Population.2022.")

# sorts df1 based on the population column in descending order
df1 <- df1 %>%
  arrange(desc(population))

# getting all data from 2020
df2020 <- df1 %>%
  filter(Year == 2020)

head(df2020)

# filter where it returns all rows where one value has NaN on all columns
null2020 <- df2020 %>%
  filter(rowSums(is.na(.)) > 0)

null2020$Country
num_rows <- nrow(null2020)
print(num_rows)


# Identify the most polluting country
theMostPolluting <- df2020[df2020$coTons == max(df2020$coTons, na.rm = TRUE), ]
theMostPolluting

```

Provides valuable insights into CO2 emissions in the year 2020, shedding
light on the most polluting country, United States, and exploring a
group of countries with undefined values. The accompanying visualization
enhances our understanding of comparative emissions and the contribution
of the undefined group.

```{r}
# arrange in ascending order based on coTons column
sorted2020 <- df2020 %>% arrange(coTons)
head(sorted2020)
mostPolluting <- df2020[order(-df2020$coTons), ][1, ]
head(mostPolluting)

```

### [Top 10 Countries in 2020]{.smallcaps}

Focusing on the top 10 countries with the highest CO2 emissions in the
year 2020. A subset of the dataset was created by selecting the top 10
countries with the highest CO2 emissions. It was visualized over the
course of the year using a line plot with various colors. United States
shows to be leading, while China is second and just over half of the
US's emissions.

```{r}
# Select the top 10 countries with the highest CO2 emissions in 2020
top10_2020 <- df2020 %>% arrange(desc(coTons)) %>% head(10)

# Filter the original dataset for the selected countries
topCountriesCO2 <- df1 %>% filter(Country %in% top10_2020$Country)

# Plot the data using ggplot2
ggplot(topCountriesCO2, aes(x = Year, y = coTons, color = Country)) +
  geom_line(size = 1) +
  labs(title = "CO2 Emissions of Top 10 Countries in 2020",
       x = "Year",
       y = "CO2 Tons") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)


```

Here is the start of the timeline throughout the centuries. This is
starting at the beginning of the dataset, so it's not a full century,
only spanning 50 years. Around 1760 was the start of the Industrial
Revolution, which meant an increase in CO2 emissions due to the increase
of coal to fuel the Industrial Revolution.

```{r}
# 18th century
cent_18 <- df1 %>% filter(Year >= 1750 & Year <= 1799)

# Select the top 10 countries with the highest CO2 emissions in the 19th century
big_cent_18 <- cent_18 %>% 
  group_by(Country) %>% 
  summarise(TotalCO2 = sum(coTons)) %>% 
  arrange(desc(TotalCO2)) %>% 
  head(10)

# Filter the original dataset for the selected countries in the 19th century
CO2_cent_18 <- cent_18 %>% filter(Country %in% big_cent_18$Country)

# Plot the data using ggplot2
ggplot(CO2_cent_18, aes(x = Year, y = coTons, color = Country)) +
  geom_line(size = 1) +
  labs(title = "CO2 Emissions of Top 10 Countries in the 18th Century",
       x = "Year",
       y = "CO2 Tons") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)


```

Next, we have the 19th century. Here we see a drastic change, the United
States made the top 10 and brought itself to second. Other countries
made an appearance as well, adding to varying CO2 emissions by the turn
of the century.

```{r}
# 19th century
cent_19 <- df1 %>% filter(Year >= 1800 & Year <= 1899)

# Select the top 10 countries with the highest CO2 emissions in the 19th century
big_cent_19 <- cent_19 %>% 
  group_by(Country) %>% 
  summarise(TotalCO2 = sum(coTons)) %>% 
  arrange(desc(TotalCO2)) %>% 
  head(10)

# Filter the original dataset for the selected countries in the 19th century
CO2_cent_19 <- cent_19 %>% filter(Country %in% big_cent_19$Country)

# Plot the data using ggplot2
ggplot(CO2_cent_19, aes(x = Year, y = coTons, color = Country)) +
  geom_line(size = 1) +
  labs(title = "CO2 Emissions of Top 10 Countries in the 19th Century",
       x = "Year",
       y = "CO2 Tons") + scale_y_continuous(labels = scales::comma)
```

Now we reach the 1900's, at first glance the graph seems like the other
countries didn't add much to the CO2 emissions and the US just
skyrocketed. If you look at the numbers on the y-axis, you will see that
the US is just an outlier, and the other countries have actually doubled
their CO2. Many factors could lead to this jump. One of them could be
the post World-War II baby boom, an increase in population equals an
increase in CO2 emissions.

```{r}
# 20th century
# Filter the original dataset for the 20th century
cent_20 <- df1 %>% filter(Year >= 1900 & Year <= 1999)

# Select the top 10 countries with the highest CO2 emissions in the 20th century
big_cent_20 <- cent_20 %>% 
  group_by(Country) %>% 
  summarise(TotalCO2 = sum(coTons)) %>% 
  arrange(desc(TotalCO2)) %>% 
  head(10)

# Filter the original dataset for the selected countries in the 20th century
CO2_cent_20 <- cent_20 %>% filter(Country %in% big_cent_20$Country)

# Plot the data using ggplot2
ggplot(CO2_cent_20, aes(x = Year, y = coTons, color = Country)) +
  geom_line(size = 1) +
  labs(title = "CO2 Emissions of Top 10 Countries in the 20th Century",
       x = "Year",
       y = "CO2 Tons") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)


```

While this graph only shows 20 years of the 21st century, it shows how
fast China is branching away from the other 8 leading countries and
catching up to the United States. In less than 15 years (starting from
2015), China doubled their CO2 emissions, making to almost half of what
the US has.

```{r}
# Filter the original dataset for the 21st century
cent_21 <- df1 %>% filter(Year >= 2000 & Year <= 2022)

# Select the top 10 countries with the highest CO2 emissions in the 21st century
big_cent_21 <- cent_21 %>% 
  group_by(Country) %>% 
  summarise(TotalCO2 = sum(coTons)) %>% 
  arrange(desc(TotalCO2)) %>% 
  head(10)

# Filter the original dataset for the selected countries in the 21st century
CO2_cent_21 <- cent_21 %>% filter(Country %in% big_cent_21$Country)

# Plot the data using ggplot2
ggplot(CO2_cent_21, aes(x = Year, y = coTons, color = Country)) +
  geom_line(size = 1) +
  labs(title = "CO2 Emissions of Top 10 Countries in the 21st Century",
       x = "Year",
       y = "CO2 Tons") +
  theme_minimal() + scale_y_continuous(labels = scales::comma)


```

### [CO2 Emissions per Area]{.smallcaps}

This graph shows the CO2 emissions per area for the top countries in
2020. Leading is the United Kingdom, their high figure suggests a
concentrated impact on the environment despite its relatively smaller
landmass. While on the opposite end is Canada, despite its expansive
landmass, they produce less CO2 per Area than smaller countries.

```{r}
area2020 = top10_2020
area2020$coByArea <- area2020$coTons / area2020$Area
area2020

ggplot(area2020, aes(x = reorder(Country, -coByArea), y = coByArea, fill = Country)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  theme_minimal() +
  labs(x = "Country", y = "CO2 Tons per Area") +
  ggtitle("CO2 Emissions per Area for Top Countries in 2020") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set3") + scale_y_continuous(labels = scales::comma)


```

### [**Developing vs Developed Countries**]{.smallcaps}

```{r}

gdp_threshold <- 25000

df_categorized <- wdi_data %>%
  mutate(GDP_per_capita = NY.GDP.MKTP.KD / SP.POP.TOTL)


df_categorized <- df_categorized %>%
  mutate(development_status = ifelse(GDP_per_capita >= gdp_threshold, "Developed", "Developing"))

df_categorized <- na.omit(df_categorized)

developed_countries <- df_categorized %>%
  filter(development_status == "Developed")

developing_countries <- df_categorized %>%
  filter(development_status == "Developing")

```

```{r}
emissions_2020 <- df_categorized %>%
  filter(Year == 2020)

top_developed <- emissions_2020 %>%
  filter(development_status == "Developed") %>%
  top_n(10, wt = CO2.emission..Tons.)

top_developing <- emissions_2020 %>%
  filter(development_status == "Developing") %>%
  top_n(10, wt = CO2.emission..Tons.)

# Plot emissions for the top 10 developed countries in 2020
developed_plot <- ggplot(top_developed, aes(x = reorder(Country, -CO2.emission..Tons.), y = CO2.emission..Tons., fill = development_status)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 CO2 Emissions in 2020: Developed Countries",
       x = "Country",
       y = "CO2 Emissions (Tons)",
       fill = "Development Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

# Plot emissions for the top 10 developing countries in 2020
developing_plot <- ggplot(top_developing, aes(x = reorder(Country, -CO2.emission..Tons.), y = CO2.emission..Tons., fill = development_status)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 CO2 Emissions in 2020: Developing Countries",
       x = "Country",
       y = "CO2 Emissions (Tons)",
       fill = "Development Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

print(developed_plot)
print(developing_plot)
```

*India is 2nd to China in terms of CO2 emissions. As more companies
attempt to gain a foothold in India for setting up manufacturing plants,
it would be interesting to see in the future whether India's CO2
emissions rise as China's falls*

```{r}
emissions_2020_excluded <- anti_join(emissions_2020, top_developed, by = c("Country", "development_status"))
emissions_2020_excluded <- anti_join(emissions_2020_excluded, top_developing, by = c("Country", "development_status"))

# Scatter plot of GDP (Per Capita) vs. CO2 emissions for 2020 (excluding top 10)
ggplot(emissions_2020_excluded, aes(x = GDP_per_capita, y = CO2.emission..Tons., color = development_status)) +
  geom_point() +
  labs(title = "GDP (Per Capita) vs. CO2 Emissions in 2020 (Excluding Top 10):\nDeveloped vs. Developing Countries",
       x = "GDP (Per Capita)",
       y = "CO2 Emissions (Tons)",
       color = "Development Status")

```

*Analyzing the relationship between GDP (Per Capita) and CO2 emissions
while distinguishing between developing and developed countries offers
insights into the impact of climate agreements. Traditionally, nations
with higher GDP have often taken the lead in establishing and adhering
to climate goals, a trend that aligns with the observations depicted in
the plot.*

## [[**Conclusion**]{.underline}]{.smallcaps}

Our analysis provides valuable insights into the changing landscape of
CO2 emissions. We observed historical emission patterns, the impact of
international agreements, and the contributions of specific countries.
Understanding these trends is crucial for policymakers and stakeholders
working towards sustainable environmental practices.

[[**Historical Emission Patterns**]{.underline}]{.smallcaps}

Our exploration of historical emission patterns provided a comprehensive
view of how different countries have contributed to the cumulative CO2
emissions since 1750. By identifying the starting points of recorded
emissions for each country, we gained insights into the trajectories of
their carbon footprints over time. This historical context aids in
understanding the varying roles nations have played in the global
emissions landscape.

[[**Impact of International Agreements**]{.underline}]{.smallcaps}

Our analysis delved into the impact of international agreements on CO2
emissions, specifically focusing on countries with populations exceeding
75 million. Visualizing emissions trends alongside the dates of major
agreements revealed nuanced shifts in emission patterns following
international efforts. This aspect of our study sheds light on the
effectiveness of collaborative initiatives in curbing and fostering a
more sustainable future.

[[**Significance for Policymakers and
Stakeholders**]{.underline}]{.smallcaps}

The insights derived from our analysis carry significant implications
for policymakers and stakeholders dedicated to environmental
sustainability. Policymakers can leverage this information to design
tailored strategies for emission reduction, accounting for historical
contributions and the impact of international collaborations.
Stakeholders, including environmental organizations and businesses, can
use these findings to align their efforts with global objectives and
contribute to a more sustainable and resilient future.

[[**Future Considerations**]{.underline}]{.smallcaps}

As we progress towards a more environmentally conscious future,
continued monitoring and analysis of CO2 emissions remain imperative.
Ongoing assessments will enable us to evaluate the effectiveness of
current policies, identify emerging trends, and adapt strategies to
address evolving challenges. Additionally, exploring the intersection of
CO2 emissions with socioeconomic factors, technological advancements,
and policy implementations can provide a holistic understanding of the
forces shaping our environmental trajectory.

In conclusion, our analysis not only illuminates the historical context
of CO2 emissions but also offers a roadmap for informed decision-making
in the pursuit of a sustainable and resilient global ecosystem. As we
collectively strive for a greener future, the lessons drawn from this
study can serve as valuable guides for shaping impactful policies and
fostering a shared commitment to environmental stewardship.
