# DS202_Final
Repo for DS202 final project

```{r}
library(dplyr)
library(tidyverse)
```

The data we are using is the CO2 emissions by country dataset. This data set
tracked CO2 emission from different countries in the world from 1750 until 2022.

```{r}
df <- read.csv("CO2 emission by countries.csv")
head(df)
length(df)
nrow(df)
summary(df)
```
There are 9 columns- country, code, calling code, year, Co2 emissions in tons,
population as of 2022, area, percent of world, and density in km2.

Country is the name of the country. Code is the country's abbreviation. Calling
code is the area code in a phone number. Year is the year the data was observed.
CO2 emissions in tons is the recorded amount of CO2 emissions by the country for
that year. Population is the number of people living in the country. Area is the 
size of the country in kilometers squared. Percent of world is how much of the
Earth the country takes up. Density in km2 is the density according to the area.

There are 59620 rows of data.

```{r}
drop_na(df)
```
